Interactive Visualization of Large Bipartite Networks Assisted by Multilevel Strategies\thanks{Supported by FAPESP, project 2017/05838-3}

### 2019-05-28 paper.tex
arrumacoes minimas

\section{Method description}\label{des}
\subsection{Fundamental concepts}\label{bac}
A bipartite network $G=(V,E)$ consists in a set $V$ of nodes which is partitioned in two subsets
with no links between nodes in the same set, i.e. $\exists\; V_1, V_2:\; V_1\cup V_2 = V,\;V_1\cap V_2 = \emptyset$, and $E\subseteq V_1 \times V_2$.
The variable names are borrowed from more traditional
Graph (network) theory, with Vertices (nodes) and Edges (links).
One may regard the network as $G=(V,E,\sigma,\omega)$, with
$\sigma\,:\,V\rightarrow \mathbb{R}^*$ and $\omega\,:\,E \rightarrow \mathbb{R}^*$,
where $\sigma(v)$ is the weight of the node $v$ and
$\omega(u,v)$ is the weight of the link $(u,v)$.

@@

\subsection{Multilevel coarsening of bipartite networks}\label{bac}
Before~\cite{alan2}, multilevel approaches were not directly applicable to bipartite networks.
Such work introduced novel and efficient matching and coarsening algorithms,
and scrutinized their validity for solving optimization problems, dimensionality reduction,
and in the preservation of essential topological features of the original network.
We here delineate such procedures in the context of our network visualization contribution.

@@

\begin{figure}[!h]\centering
 \includegraphics[width=\textwidth]{mlf}
  \caption{Phases of the multilevel optimization.
  A sequence of incrementally reduced networks is obtained in the coarsening phase.
  An initial solution found at the coarsest network, in the solution finding phase,
  and is projected back to the initial, uncoarsened, network, in the uncoarsening phase.
  Further details are given in Section~\ref{bac} and specificities related to these
  phases when applied to interactive data visualization are in Sections~\ref{net} and~\ref{nav}.
  }\label{mlf}
\end{figure}

@@

\begin{figure}[!h]\centering
 \includegraphics[width=0.8\textwidth]{fnav___}
  \caption{The navigation pathway of bipartite networks obtained through a multilevel strategy
  and auxiliary tools.
  Compliant with the Shneiderman's mantra, first and overview of the bipartite network is provided.
  Then the user may select an specific portion of the network to visualize in further detail,
  apply filters and visual mappings of topological features, and request details.
  These operations may be carried on in any order and any number of times the analyst understand convenient.
  }\label{fnav}
\end{figure}

@@

\begin{figure}[!h]\centering
 \includegraphics[width=\textwidth]{overall____}
  \caption{The MlBiNetVis interface. (1) is the dropdown menu for choosing the network to analyse or uploading new ones.
  The gray box in (2) contains widgets to select and parametrize the coarsening algorithm.
  (3) is the render button, which maps the network to the screen.
  In (4) are auxiliary widgets for selecting the  layout, show/hide of links, and for the interface to output information requested by the user or useful to assist navigation.
  In (5) is an interactive table which holds information about the levels and layers of the bipartite network and may be used to set the visualization on the canvas.
  In (6) is the toolbar that enables the navigation of the network and fine-tuning the visualization.
  In (7) is the canvas in which the network is drawn through node-link diagrams.
  Section~\ref{use} describes the use of these elements in further detail.
  }\label{fpage0}
\end{figure}

\subsection{The use of MlBiNetVis}\label{use}
% network upload, format.
% layouts available
% algorithms available for coarsening, and their parametrization
% initialization of the visualization: click on render
% tools available and their operation
In using MlBiNetVis,
the analyst first selects a network of interest to be explored.
One may also upload a network and then select it on the same drop-down menu.
The user may then wish to select one of the available network layout
algorithms and tune the coarsening to be performed, but default settings are reasonable
for the newcomer.
Upon request given by hitting the ``render network'' button,
the network is mapped to the screen.
Subsequent usage relies on manipulation of the visual mapping presented on the
canvas by means of actions
performed through managing the tools and related mouse operations.
These steps are illustrated in Figure~\ref{initial}.
\begin{figure}[!h]\centering
 \includegraphics[width=\textwidth]{initial_}
  \caption{The elements in the initial step of the MlBiNetVis usage.
  The button at the top is for network upload and selection.
  In the gray box, the user chooses and sets the coarsening algorithm as detailed in Appendix~\ref{ap}.
  The blue button at the bottom is a drop-down menu where the user picks the node-link
  layout algorithm.
  The green button is for rendering the network to the canvas and initializing further widgets.
  By hitting the green button, all the elements in this image is made disabled, except for the layout
  button.
  Further information is given in Section~\ref{use}.
  }\label{initial}
\end{figure}

@@

\begin{figure}[!h]\centering
 \includegraphics[width=\textwidth]{secondPhase}
  \caption{The elements in second step of the MlBiNetVis usage, dedicated to navigation on the network and transformation of the visualization.
  The user may change the layout used when supernodes are opened and the predecessors are then shown on the screen.
  Links may be shown or hided for all the levels by using the checkbox.
  The text area assists in the output of information requested or useful for navigation.
  The canvas holds the node-link representations of the levels, which are transformed using the tools in the toolbar and the table in the right-side.
  Such table also makes evident key information about the status of the visualization.
  Further information is given in Section~\ref{use}.
  }\label{secondPhase}
\end{figure}

@@

\bibliographystyle{splncs04}
\bibliography{biblio}
%
% \begin{thebibliography}{8}
% \bibitem{ref_article1}
% Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
% 
% \bibitem{ref_lncs1}
% Author, F., Author, S.: Title of a proceedings paper. In: Editor,
% F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
% Springer, Heidelberg (2016). \doi{10.10007/1234567890}
% 
% \bibitem{ref_book1}
% Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
% Location (1999)
% 
% \bibitem{ref_proc1}
% Author, A.-B.: Contribution title. In: 9th International Proceedings
% on Proceedings, pp. 1--2. Publisher, Location (2010)
% 
% \bibitem{ref_url1}
% LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
% Oct 2017
% \end{thebibliography}
\end{document}

>>>>

\section{Method description}\label{des}
\subsection{Fundamental concepts}
A bipartite network $G=(V,E)$ consists in a set $V$ of nodes which is partitioned in two subsets
with no links between nodes in the same set, i.e. $\exists\; V_1, V_2:\; V_1\cup V_2 = V,\;V_1\cap V_2 = \emptyset$, and $E\subseteq V_1 \times V_2$.
The variable names are borrowed from more traditional
Graph (network) theory, with Vertices (nodes) and Edges (links).
One may regard the network as $G=(V,E,\sigma,\omega)$, with
$\sigma\,:\,V\rightarrow \mathbb{R}^*$ and $\omega\,:\,E \rightarrow \mathbb{R}^*$,
where $\sigma(v)$ is the weight of the node $v$ and
$\omega(u,v)$ is the weight of the link $(u,v)$.

@@

\subsection{Multilevel coarsening of bipartite networks}
Before~\cite{alan2}, multilevel approaches were not directly applicable to bipartite networks.
Such work introduced novel and efficient matching and coarsening algorithms,
and scrutinized their validity for solving optimization problems, dimensionality reduction,
and in the preservation of essential topological features of the original network.
We here delineate such procedures in the context of our network visualization contribution.

@@

\begin{figure}\centering
 \includegraphics[width=\textwidth]{mlf}
  \caption{Phases of the multilevel optimization.
  A sequence of incrementally reduced networks is obtained in the coarsening phase.
  An initial solution found at the coarsest network, in the solution finding phase,
  and is projected back to the initial, uncoarsened, network, in the uncoarsening phase.
  Further details are given in Section~\ref{des} and specificities related to these
  phases when applied to interactive data visualization are in Sections~\ref{net} and~\ref{nav}.
  }\label{mlf}
\end{figure}

@@

\begin{figure}\centering
 \includegraphics[width=0.8\textwidth]{fnav___}
  \caption{The navigation pathway of bipartite networks obtained through a multilevel strategy
  and auxiliary tools.
  Compliant with the Shneiderman's mantra, first and overview of the bipartite network is provided.
  Then the user may select an specific portion of the network to visualize in further detail,
  apply filters and visual mappings of topological features, and request details.
  These operations may be carried on in any order and any number of times the analyst understand convenient.
  }\label{fnav}
\end{figure}

@@

\begin{figure}\centering
 \includegraphics[width=\textwidth]{overall____}
  \caption{The MlBiNetVis interface. (1) is the dropdown menu for choosing the network to analyse or uploading new ones.
  The gray box in (2) contains widgets to select and parametrize the coarsening algorithm.
  (3) is the render button, which maps the network to the screen.
  In (4) are auxiliary widgets for selecting the  layout, show/hide of links, and for the interface to output information requested by the user or useful to assist navigation.
  In (5) is an interactive table which holds information about the levels and layers of the bipartite network and may be used to set the visualization on the canvas.
  In (6) is the toolbar that enables the navigation of the network and fine-tuning the visualization.
  In (7) is the canvas in which the network is drawn through node-link diagrams.
  Section~\ref{use} describes the use of these elements in further detail.
  }\label{fpage0}
\end{figure}

\subsection{The use of MlBiNetVis}\label{use}
% network upload, format.
% layouts available
% algorithms available for coarsening, and their parametrization
% initialization of the visualization: click on render
% tools available and their operation
In using MlBiNetVis,
the analyst first selects a network of interest to be explored.
One may also upload a network and then select it on the same drop-down menu.
The user may then wish to select one of the available network layout
algorithms and tune the coarsening to be performed, but default settings are reasonable
for the newcomer.
Upon request given by hitting the ``render network'' button,
the network is mapped to the screen.
Subsequent usage relies on manipulation of the visual mapping presented on the
canvas by means of actions
performed through managing the tools and related mouse operations.
These steps are illustrated in Figure~\ref{initial}.
\begin{figure}\centering
 \includegraphics[width=\textwidth]{initial_}
  \caption{The elements in the initial step of the MlBiNetVis usage.
  The button at the top is for network upload and selection.
  In the gray box, the user chooses and sets the coarsening algorithm as detailed in Appendix~\ref{ap}.
  The blue button at the bottom is a drop-down menu where the user picks the node-link
  layout algorithm.
  The green button is for rendering the network to the canvas and initializing further widgets.
  By hitting the green button, all the elements in this image is made disabled, except for the layout
  button.
  Further information is given in Section~\ref{use}.
  }\label{initial}
\end{figure}

@@

\begin{figure}\centering
 \includegraphics[width=\textwidth]{secondPhase}
  \caption{The elements in second step of the MlBiNetVis usage, dedicated to navigation on the network and transformation of the visualization.
  The user may change the layout used when supernodes are opened and the predecessors are then shown on the screen.
  Links may be shown or hided for all the levels by using the checkbox.
  The text area assists in the output of information requested or useful for navigation.
  The canvas holds the node-link representations of the levels, which are transformed using the tools in the toolbar and the table in the right-side.
  Such table also makes evident key information about the status of the visualization.
  Further information is given in Section~\ref{use}.
  }\label{secondPhase}
\end{figure}

@@

\bibliographystyle{splncs04}
\bibliography{biblio}
